Generalized Tractability

### 2023-04-02 GeneralizedTractability_230310.tex
Fred M2 2023 Mac Updates

\subsection{Notation and fundamental definitions}
Let $T$ be a tractability function defined as follows,
\[
T :(0,\infty) \times \mathbb{N} \times [0,\infty)^s \rightarrow (0,\infty),
\]
\fred{I think that we do not need to bound $T$ away from $0$.}
A tractability function is a function of simple form that provides an upper bound on the computational complexity of a problem.  Some examples of tractability functions commonly considered are provided in Section \ref{sec:examples}.

@@

Define
\[
j (\varepsilon,\vp):= \lfloor C_{\vp}\, T(\varepsilon^{-1},1,\vp)\rfloor +1, \quad
j (\infty,\vp) = \lim_{\varepsilon \to \infty} j (\varepsilon,\vp) \ge 1 .
\]
Thus, it follows by \eqref{eq:lambda_K_strong3} that $\lambda_{j(\varepsilon,\vp),d} \le \varepsilon$.
Note furthermore that we always have
\[
j(\varepsilon,\vp)\le C_{\vp}\, T(\varepsilon^{-1},1,\vp)+1 \le C_{\vp} T(\lambda_{j(\varepsilon,\vp),d}^{-1},1,\vp)+1, \qquad \forall \varepsilon > 0,
\]
since
$T(\cdot,1, \vp)$ is non-decreasing.

For $\varepsilon$ taking on all positive values, $j(\varepsilon,\vp)$ takes on all values greater than or equal to $j(\infty,\vp)$, so
\[
j\le C_{\vp} T(\lambda_{j,d}^{-1},1,\vp)+1, \qquad \forall j \ge j(\infty,\vp).
\]
This implies via our technical assumption \eqref{eq:ptauassume} that
\begin{gather*}
 K_{\vp,\tau}\,T (\lambda_{j,d}^{-1},1,\tau \vp) \ge
 [T(\lambda_{j,d}^{-1},1, \vp)]^\tau
 \ge
  \left[\frac{(j-1)}{C_{\vp}}\right]^\tau \qquad \forall j \ge j(\infty,\vp) \\
 \frac{1}{T (\lambda_{j,d}^{-1},1,\tau \vp)} \le
\frac{K_{\vp,\tau}\, C_{\vp}^\tau}{(j-1)^\tau}\qquad \forall j \ge j(\infty,\vp)
\end{gather*}
Summing both sides of this last inequality for $j$ from $j(\infty,\vp) \ge 1$ to $\infty$ yields 
\begin{align*}
\sup_{d\in\naturals} \sum_{j=j (\infty,\vp)}^\infty \frac{1}{T (\lambda_{j,d}^{-1},1, \tau \vp)}
& \le  \sup_{d\in\naturals} \frac{1}{T (\lambda_{1,d}^{-1},1, \tau \vp)} + K_{\vp,\tau}\, C_{\vp}^\tau
\sum_{j=2}^\infty \frac{1}{(j-1)^\tau} \\
& \le\frac{1}{T (\lambda_{\max}^{-1},1, \tau \vp)} +  K_{\vp,\tau}\, C_{\vp}^\tau
\zeta (\tau)  < \infty,
\end{align*}
where $\zeta$ denotes the Riemann zeta function.
This yields \eqref{eq:strong_tractiff3} with $\vp$ replaced by $\vp' = \tau \vp$ and $L_{\vp'} = j (\infty,\vp)$, and so we see the necessity of $\eqref{eq:strong_tractiff3}$.

@@

Let $J(\vp,d): = \lceil L_{\vp}\, T(0,d,\vp)\rceil$.  Then this upper bound on the complexity can be further bounded above by
\begin{align*}
       \comp(\varepsilon,d)
       & = T(\varepsilon^{-1},d,\vp) \Biggl [ \sum_{i=1}^{J(\vp,d) -1} \frac{1}{T(\lambda_{i, d}^{-1},d,\vp)}
       + \underbrace{\sum_{i=J(\vp,d)}^\infty \frac{1}{T(\lambda_{i, d}^{-1},d,\vp)}}_{\le S_{\vp} \text{ by \eqref{eq:tractiff4}}} \Biggr ] \\
       & \le T(\varepsilon^{-1},d,\vp)\left[ \frac{J(\vp,d)-1}{T(\lambda_{\max}^{-1},d,\vp)} + S_{\vp}\right]
        \\
       & \le T(\varepsilon^{-1},d,\vp)\left[\frac{L_{\vp}\, T(0,d,\vp)}{T(\lambda_{\max}^{-1},d,\vp)} + S_{\vp}\right] \\
       & \le\frac{L_{\vp}}{T(\lambda_{\max}^{-1},d,\vp)} \,T(\varepsilon^{-1},d,\vp) T(0,d,\vp) + S_{\vp}\, T(\varepsilon^{-1},d,\vp)\\
       & \le\frac{L_{\vp}}{T(\lambda_{\max}^{-1},d,\vp)} \,T(\varepsilon^{-1},d,\vp) T(\varepsilon^{-1},d,\vp) + S_{\vp}\, T(\varepsilon^{-1},d,\vp)\\
       & \le \frac{L_{\vp}\,K_{\vp,2}}{T(\lambda_{\max}^{-1},d,\vp)}  T(\varepsilon^{-1},d,2\vp)  + S_{\vp}\, T(\varepsilon^{-1},d,\vp)  \qquad \text{by \eqref{eq:ptauassume}}\\
       & \le \underbrace{\left[\frac{L_{\vp}\,K_{\vp,2}}{T(\lambda_{\max}^{-1},d,\vp)} + S_{\vp}\right]}_{=:C_{2\vp}} \,T(\varepsilon^{-1},d,2\vp)
\end{align*}
It follows that we have $T$-tractability, which shows sufficiency of \eqref{eq:tractiff4}. \\

\noindent \textbf{Necessary condition:}\\
Suppose that we have
$T$-tractability. That is, for some $\vp \ge \vzero$, there exists a positive constant $C_{\vp}$ such that
\[
\comp(\varepsilon,d)\le C_{\vp}\, T(\varepsilon^{-1},d,\vp) \qquad \forall \varepsilon > 0,  \ d \in \naturals.
\]
Since the eigenvalues $\lambda_{j,d}$ are non-increasing, we have
\begin{equation}\label{eq:lambda_K4}
	\lambda_{\lfloor C_{\vp}\, T(\varepsilon^{-1},d,\vp)\rfloor +1,d}\le \varepsilon.
\end{equation}
Define
\[
j (\varepsilon,d,\vp):= \lfloor C_{\vp}\, T(\varepsilon^{-1},d,\vp)\rfloor +1, \quad
j (\infty,d,\vp) = \lfloor C_{\vp}\, T(0,d,\vp)\rfloor +1.
\]
Thus, it follows by \eqref{eq:lambda_K4} that $\lambda_{j(\varepsilon,d,\vp),d} \le \varepsilon$.
Note furthermore that we always have
\[
j(\varepsilon,d,\vp)\le C_{\vp}\, T(\varepsilon^{-1},1,\vp)+1 \le C_{\vp} T(\lambda_{j(\varepsilon,d,\vp),d}^{-1},1,\vp)+1, \qquad \forall \varepsilon > 0, \ d \in \naturals
\]
since
$T(\cdot,d,\vp)$ is non-decreasing.

For $\varepsilon$ taking on all positive values, $j(\varepsilon,d,\vp)$ takes on all values greater than or equal to $j(\infty,d,\vp)$, so
\[
j\le C_{\vp}\, T(\varepsilon^{-1},d,\vp)+1 \le C_{\vp} T(\lambda_{j,d}^{-1},d,\vp)+1, \qquad \forall j \ge j(\infty,d,\vp).
\]
This implies via our technical assumption \eqref{eq:ptauassume} that
\begin{equation*}
	K_{\vp,\tau}\,T (\lambda_{j,d}^{-1},d,\tau \vp) \ge
	[T(\lambda_{j,d}^{-1},d, \vp)]^\tau
	\ge
	\left[\frac{(j-1)}{C_{\vp}}\right]^\tau \qquad \forall j \ge j(\infty,d,\vp)
\end{equation*}
Summing both sides of the inequality yields and noting that $L_{\vp} \le L(\tau\vp)$, it follows that
\peter{(Peter: Why does $L(\cdot)$ ocur here?)} \fred{Because I got lost a bit.}

\begin{multline*}
	\sup_{d\in\naturals} \sum_{j=j(\infty,d,\vp)}^\infty \frac{1}{T (\lambda_{j,d}^{-1},d, \tau \vp)} \le
	\sup_{d\in\naturals} \sum_{j=j(\infty,d,\vp)}^\infty \frac{1}{T (\lambda_{j,d}^{-1},d, \tau \vp)} \\
	\le K_{\vp,\tau}\, C_{\vp}^\tau
	\sum_{j=j(\infty,d,\vp)}^\infty \frac{1}{(j-1)^\tau} \le
	K_{\vp,\tau}\, C_{\vp}^\tau\, \zeta (\tau) < \infty,
\end{multline*}
where $\zeta$ denotes the Riemann zeta function.

Note now that, \peter{for any $d\in\naturals$,}
\[
 j(\infty,d,\vp) = \lfloor C_{\vp}\, T(0,d,\vp)\rfloor +1 \le \lfloor C_{\vp}^\tau\, T(0,d,\tau \vp)\rfloor +1.
\]
So, there exists $L(\tau \vp)$ such that
\[
\peter{\sup_{d\in\naturals}}\,\, \sum_{j=\lceil L(\tau \vp) T(0,d,\tau \vp)\rceil}^\infty \frac{1}{T(\lambda_{j,d}^{-1},d, \tau\vp)}<\infty.
\]
\fred{Fred:  Not sure how $C_{\vp}^\tau$ becomes $L(\tau \vp)$?}

\peter{Peter: You are right. We could probably solve this by just saying above that
\[
 j(\infty,d,\vp) = \lfloor C_{\vp}\, T(0,d,\vp)\rfloor +1 \le \lfloor C_{\vp}\, T(0,d,\tau \vp)\rfloor +1,
\]
and then introduce $L(\tau \vp)$ to bound $C_{\vp}$ and also incorporate the ``+1''?
}

This yields \eqref{eq:tractiff4} with $\vp$ replaced by $\tau \vp$, and so we see the necessity of $\eqref{eq:tractiff4}$.

>>>>

\subsection{Notation and fundamental definitions}
Let $T$ be a tractability function defined as follows,
\[
T :(0,\infty) \times \mathbb{N} \times [0,\infty)^s \rightarrow (0,\infty),
\]
\fred{I think that we do not need to bound $T$ away from $0$, but doing so may help satisfy \eqref{eq:ptauassume}.}
A tractability function is a function of simple form that provides an upper bound on the computational complexity of a problem.  Some examples of tractability functions commonly considered are provided in Section \ref{sec:examples}.

@@

Define
\[
i(\varepsilon,\vp):= \lfloor C_{\vp}\, T(\varepsilon^{-1},1,\vp)\rfloor +1, \quad
i(\infty,\vp) = \lim_{\varepsilon \to \infty}  i(\varepsilon,\vp) \ge 1 .
\]
Thus, it follows by \eqref{eq:lambda_K_strong3} that $\lambda_{i(\varepsilon,\vp),d} \le \varepsilon$.
Note furthermore that we always have
\[
i(\varepsilon,\vp)\le C_{\vp}\, T(\varepsilon^{-1},1,\vp)+1 \le C_{\vp} T(\lambda_{i(\varepsilon,\vp),d}^{-1},1,\vp)+1, \qquad \forall \varepsilon > 0,
\]
since
$T(\cdot,1, \vp)$ is non-decreasing.

For $\varepsilon$ taking on all positive values, $i(\varepsilon,\vp)$ takes on all values greater than or equal to $i(\infty,\vp)$, so
\[
i\le C_{\vp} T(\lambda_{i,d}^{-1},1,\vp)+1, \qquad \forall i \ge i(\infty,\vp).
\]
This implies via our technical assumption \eqref{eq:ptauassume} that
\begin{gather*}
 K_{\vp,\tau}\,T (\lambda_{i,d}^{-1},1,\tau \vp) \ge
 [T(\lambda_{i,d}^{-1},1, \vp)]^\tau
 \ge
  \left[\frac{(i-1)}{C_{\vp}}\right]^\tau \qquad \forall i \ge i(\infty,\vp) \\
 \frac{1}{T (\lambda_{i,d}^{-1},1,\tau \vp)} \le
\frac{K_{\vp,\tau}\, C_{\vp}^\tau}{(i-1)^\tau}\qquad \forall i \ge i(\infty,\vp).
\end{gather*}
Summing both sides of this last inequality over $i$ from $i(\infty,\vp) +1 \ge 2$ to $\vinfty$ yields
\begin{equation*}
\sup_{d\in\naturals} \sum_{i=i(\infty,\vp) +1 }^\infty \frac{1}{T (\lambda_{i,d}^{-1},1, \tau \vp)}
 \le  K_{\vp,\tau}\, C_{\vp}^\tau
\sum_{i=2}^\infty \frac{1}{(i-1)^\tau} \\
 \le K_{\vp,\tau}\, C_{\vp}^\tau
\zeta (\tau)  < \infty,
\end{equation*}
where $\zeta$ denotes the Riemann zeta function.
This yields \eqref{eq:strong_tractiff3} with $\vp$ replaced by $\vp' = \tau \vp$ and $L_{\vp'} = i(\infty,\vp) +1$, and so we see the necessity of $\eqref{eq:strong_tractiff3}$.

@@

Let $J(\vp,d): = \lceil L_{\vp}\, T(0,d,\vp)\rceil$.  Then this upper bound on the complexity can be further bounded above by
\begin{align*}
       \comp(\varepsilon,d)
       & = T(\varepsilon^{-1},d,\vp) \Biggl [ \sum_{i=1}^{J(\vp,d) -1} \frac{1}{T(\lambda_{i, d}^{-1},d,\vp)}
       + \underbrace{\sum_{i=J(\vp,d)}^\infty \frac{1}{T(\lambda_{i, d}^{-1},d,\vp)}}_{\le S_{\vp} \text{ by \eqref{eq:tractiff4}}} \Biggr ] \\
       & \le T(\varepsilon^{-1},d,\vp)\left[ \frac{J(\vp,d)-1}{T(\lambda_{\max}^{-1},d,\vp)} + S_{\vp}\right]
        \\
       & \le T(\varepsilon^{-1},d,\vp)\left[\frac{L_{\vp}\, T(0,d,\vp)}{T(\lambda_{\max}^{-1},1,\vp)} + S_{\vp}\right] \\
       & \le\frac{L_{\vp}}{T(\lambda_{\max}^{-1},1,\vp)} \,T(\varepsilon^{-1},d,\vp) T(0,d,\vp) + S_{\vp}\, T(\varepsilon^{-1},d,\vp)\\
       & \le\frac{L_{\vp}}{T(\lambda_{\max}^{-1},1,\vp)} \,T(\varepsilon^{-1},d,\vp) T(\varepsilon^{-1},d,\vp) + S_{\vp}\, T(\varepsilon^{-1},d,\vp)\\
       & \le \frac{L_{\vp}\,K_{\vp,2}}{T(\lambda_{\max}^{-1},1,\vp)}  T(\varepsilon^{-1},d,2\vp)  + S_{\vp}\, T(\varepsilon^{-1},d,\vp)  \qquad \text{by \eqref{eq:ptauassume}}\\
       & \le \underbrace{\left[\frac{L_{\vp}\,K_{\vp,2}}{T(\lambda_{\max}^{-1},1,\vp)} + S_{\vp}\right]}_{=:C_{2\vp}} \,T(\varepsilon^{-1},d,2\vp)
\end{align*}
It follows that we have $T$-tractability, which shows sufficiency of \eqref{eq:tractiff4}. \\

\noindent \textbf{Necessary condition:}\\
Suppose that we have
$T$-tractability. That is, for some $\vp \ge \vzero$, there exists a positive constant $C_{\vp}$ such that
\[
\comp(\varepsilon,d)\le C_{\vp}\, T(\varepsilon^{-1},d,\vp) \qquad \forall \varepsilon > 0,  \ d \in \naturals.
\]
Since the sequence of eigenvalues $\lambda_{1,d}, \lambda_{2,d}, \ldots $ is non-increasing, we have
\begin{equation}\label{eq:lambda_K4}
	\lambda_{\lfloor C_{\vp}\, T(\varepsilon^{-1},d,\vp)\rfloor +1,d}\le \varepsilon \qquad \forall \varepsilon > 0,  \ d \in \naturals..
\end{equation}
Define the integers
\[
i (\varepsilon,d,\vp):= \lfloor C_{\vp}\, T(\varepsilon^{-1},d,\vp)\rfloor +1, \qquad
i (\infty,d,\vp) = \lfloor C_{\vp}\, T(0,d,\vp)\rfloor +1 \ge 1.
\]
Thus, it follows by \eqref{eq:lambda_K4} that $\lambda_{i(\varepsilon,d,\vp),d} \le \varepsilon$.
Note furthermore that we always have
\[
i(\varepsilon,d,\vp)\le C_{\vp}\, T(\varepsilon^{-1},d,\vp)+1 \le C_{\vp} T(\lambda_{i(\varepsilon,d,\vp),d}^{-1},d,\vp)+1, \qquad \forall \varepsilon > 0, \ d \in \naturals,
\]
since
$T(\cdot,d,\vp)$ is non-decreasing.

For $\varepsilon$ taking on all positive values, $i(\varepsilon,d,\vp)$ takes on all values greater than or equal to $i(\infty,d,\vp)$, so
\[
i\le C_{\vp}\, T(\varepsilon^{-1},d,\vp)+1 \le C_{\vp} T(\lambda_{i,d}^{-1},d,\vp)+1, \qquad \forall i \ge i(\infty,d,\vp).
\]
This implies via our technical assumption \eqref{eq:ptauassume} that for any $\tau > 1$,
\begin{gather*}
	K_{\vp,\tau}\,T (\lambda_{i,d}^{-1},d,\tau \vp) \ge
	[T(\lambda_{i,d}^{-1},d, \vp)]^\tau
	\ge
	\left[\frac{(i-1)}{C_{\vp}}\right]^\tau \qquad \forall i \ge i(\infty,d,\vp), \\
	 \frac{1}{T (\lambda_{i,d}^{-1},d,\tau \vp)} \le
	\frac{K_{\vp,\tau}\, C_{\vp}^\tau}{(i-1)^\tau}\qquad \forall i \ge i(\infty,\vp).
\end{gather*}

Let $\vp' = \tau \vp$.  Note that $i(\infty,d,\cdot)$ is nondecreasing.  Summing both sides of the inequality over $i$ from $i(\infty,d,\vp') + 1 \ge i(\infty,d, \vp) + 1 \ge 2$ to $\infty$ it follows that
\begin{equation*}
	\sup_{d\in\naturals} \, \, \sum_{i=i (\infty,d,\vp')+1}^\infty \frac{1}{T (\lambda_{i,d}^{-1},d,\vp')}
	 \le  K_{\vp,\tau}\, C_{\vp}^\tau
	\sum_{i=2}^\infty \frac{1}{(i-1)^\tau} \\
	 \le  K_{\vp,\tau}\, C_{\vp}^\tau
	\zeta (\tau)  < \infty,
\end{equation*}
where $\zeta$ denotes the Riemann zeta function.

Note now that, for any $d\in\naturals$,
\begin{align*}
 i(\infty,d,\vp') & = \lfloor C_{\vp'}\, T(0,d,\vp')\rfloor +1 \le  \left[C_{\vp'} + \frac{1}{T(0,d,\vp')}   \right]\, T(0,d,\vp')\\
 & \le  \underbrace{\left[C_{\vp'} + \frac{1}{T(0,d,\vp')}   \right]}_{=:L_{\vp'}}\, T(0,d,\vp')
 \le \lceil L_{\vp'} T(0,d,\vp')\rceil .
\end{align*}
For this choice of $ L_{\vp'}$ we have
\[
\sup_{d\in\naturals}}\,\, \sum_{i=\lceil L(\vp') T(0,d,\vp')\rceil}^\infty \frac{1}{T(\lambda_{i,d}^{-1},d, \vp')}
\le
\sup_{d\in\naturals} \, \, \sum_{i=i (\infty,d,\vp')+1}^\infty \frac{1}{T (\lambda_{i,d}^{-1},d,\vp')} <
\infty.
\]
This yields \eqref{eq:tractiff4} with $\vp$ replaced by $\vp'$, so we see the necessity of $\eqref{eq:tractiff4}$.
